{\large Gravitational wave sources in our Galactic backyard:}\\{\normalsize Models for detecting BHBH, BHNS and NSNS binaries with LISA

### 2021-11-16 paper/tex/sections/results.tex
most comments up to variations resolved

In this section we present our predictions for the population of detectable LISA sources. In total we expect $124$ detections in a 4-year LISA mission, of which $\BHBHFourYear{}$, $\BHNSFourYear{}$ and $\NSNSFourYear{}$ are BHBHs, BHNSs and NSNSs respectively, based on our fiducial simulations. In the remainder of this section, we discuss where the sources are expected relative to LISA's sensitivity curve (Sec.~\ref{sec:dcos_on_sc}), their properties (Sec.~\ref{sec:fiducial_distributions}), their locations in the Milky Way (Sec.~\ref{sec:mw_detectable_distribution}), their formation channels (Sec.~\ref{sec:progenitors_and_formation}) and finally we discuss the expected SNR and how accurately we expect that the parameters can be measured (Sec.~\ref{sec:measurement_uncertainties}). Note that all results shown in this section are based on our fiducial simulations. A discussion of the impact of variations in the physics assumptions is provided in Sec.~\ref{sec:variations}.

@@

% For both the BHBHs and BHNSs, we see that the black hole mass distribution favours low masses. About $90\%$ of BHs have masses below $11 \unit{M_{\odot}}$, in our fiducial simulations shown here. 
% Unlike the current network of ground-based GW observatories, LISA will be equally sensitive to inspiraling BHBHs throughout the entire Milky Way, regardless of their mass (given that all inspiraling BHBH pass through the LISA sensitivity band \todo{citation?}). 
% Hence we expect LISA to provide us with an unbiased view of the distribution of BH masses in inspiraling compact object binaries.

For both the BHBHs and BHNSs, we see that the black hole mass distribution favours low masses. About $90\%$ of BHs have masses below $11 \unit{M_{\odot}}$, in our fiducial simulations shown here. This is in stark contrast with the masses of the BHs that have been detected so far by ground-based GW detectors, where heavy BHs, with masses of about $30\unit{M_{\odot}}$ and higher have been common. Ground-based detectors are strongly biased towards high mass systems, since they can be detected out to larger distances. Instead, LISA is, in principle, sensitive to inspiraling BHBHs throughout the entire Milky Way regardless of their mass, as long as we catch them at the right time when their dominant frequencies coincide with the range where LISA is sensitive.

A further reason that we predict that LISA will not detect high mass systems is that the population of BHs that are detectable by LISA primarily come from progenitors that formed from high metallicity gas according to our simulations. Stellar winds are stronger at high metallicity leading to increased mass loss. This affects the mass of the most massive black holes that can be formed \citep{Belczynski+2010}.

We also note that our BH mass distribution extends down below $5\unit{M_{\odot}}$ to $2.5 \unit{M_{\odot}}$ which is our fiducial maximum neutron star mass. The BHs in this simulation fill the so-called ``lower mass gap'' marked as a grey band \citep{Ozel+2010,Farr+2011}, see also \citet{Shao+2021} who recently pointed this out. This prediction is sensitive to adopted model for fall back during the SN explosion as we discuss this further in Sec.~\ref{sec:lower_mass_gap}

\paragraph{Mass Ratio} The mass ratio distributions for each DCO type are very distinct from one another, as can be seen in Fig.~\ref{fig:fiducial_pdf_distributions}c. The majority of NSNSs have a mass ratio close to unity, with $\NSNSqAbovePointEight{}$ of systems having $q > 0.8$. The reason for the concentration around equal masses is that most NSs are formed either through electron-capture supernovae (ECSN) or from low mass stars in our simulations. We assume a remnant mass for any NS formed through ECSN of $1.26 \unit{M_{\odot}}$ (see Sec.~\ref{app:fiducial_physics}). The remnant mass prescription that we use assumes a fixed fallback mass for any star with a CO core mass less than $2.5 \unit{M_\odot}$, such that many NSs end up with an identical mass of $1.278 \unit{M_\odot}$ \citep[see][Eq.~19]{Fryer+2012}. This means that many NSs are formed with equal masses and hence we see a mass ratio distribution peaked around unity.

@@

At the same time, stellar wind mass loss disfavours the formation of black holes with comparable masses. Such systems would have originated from progenitors that also started with comparable masses. The rather massive secondaries in these systems (especially after they accreted from the primary) experience very strong stellar wind mass loss due to LBV-like eruptions. This limits the amount by which they can expand \citep[e.g.][]{vanSon+2021}. At the same time, wind mass loss (and also stable and more conservative mas transfer in these systems) lead to widening of the orbit. Both effects, the reduced expansion and increased widening of the orbit, tend to prevent the secondaries from being able to fill their Roche lobe. This thus limits the number of systems that experience the reverse mass transfer or common envelope phase needed shrink the orbits and make them tight enough to be detected as gravitational wave sources.

We find that detectable BHNSs have even more unequal mass ratios. Moreover, the mass ratio distribution is bimodal, where the two peaks arise from two distinct formation scenarios. Around two thirds of detectable BHNSs experience at least one common-envelope event, whilst the last third are formed through only stable mass transfer. The first peak at $q = 0.18$ is from systems that experience at least one common-envelope and occurs at the expected mass ratio, which approximately follows the mean BH mass ($\sim 6.5 \unit{M_\odot}$) and NS mass ($\sim 1.2 \unit{M_\odot}$). Yet we also see a second peak at higher mass ratios around $q = 0.34$, which arises from the fraction of the population that underwent only stable mass transfer phases. The stability of mass transfer depends on the mass ratio, preferentially forming systems with more equal masses, i.e.\ at higher $q$.

\paragraph{Eccentricity} In Fig.~\ref{fig:fiducial_pdf_distributions}d we show the eccentricity distributions. We find that most systems (73\%) will have eccentricities larger than $0.01$ during the LISA mission, which should in principle be detectable according to \citet{Nishizawa+2016}. This means that we will potentially be able to use eccentricity to distinguish these sources from WD sources, which are expected to have little to no eccentricity (see Sec.~\ref{sec:WDWD_distinguish}). We note that several previous studies assumed all systems to be circular when calculating the detection rates \citep[e.g.][]{Liu+2014, Lamberts+2018, Sesana+2020}. See Sec.~\ref{sec:compare_studies} for a discussion of the impact of this assumption.

For systems with eccentricities higher than $e\gtrsim0.3$, most gravitational wave energy is emitted in higher harmonics. Such systems are more rare, but we find them to be significant among the BHBH population, where they account for \BHBHHighlyEccentric{} of the systems.

The BHBHs that are still eccentric when reaching the LISA band are primarily systems formed through the stable mass transfer channel. %\citep{vandenHeuvel+2017}. 
These systems are still relatively wide just prior to formation of the second BH. If the kick is oriented roughly in opposite direction to the orbital motion and has a velocity that is of similar magnitude as the orbital velocity, it will lead to the formation of a highly eccentric system.

It is rare to get such a ``lucky kick'', but there are a few effects that favour this for BHBH systems. The kicks of BHs are reduced by fallback and they are thus less likely to disrupt the system.  Moreover, because of their higher masses, BHBHs can be observed already at lower orbital frequencies. This means that they have not had as much time to circularise and so still have significant eccentricity by the time of the LISA mission. Finally, LISA favours the detection of eccentric systems, if all other properties are held fixed. This is because the gravitational wave emission is stronger (Eq.~\ref{eq:peters_f}) and the energy is emitted at higher frequencies \citep[Eq.~20]{Peters+1963} where LISA is more sensitive.

@@

\paragraph{Time since formation} 
In Fig.~\ref{fig:fiducial_pdf_distributions}e we show how long ago the LISA detectable DCOs formed. For comparison we show the star formation history we have assumed in our Milky Way model (grey dotted line). Star formation was highest at early times $6-12\unit{Gyr}$ ago, after which it declined. In contrast, the LISA detectable DCOs primarily formed in the relatively \textit{recent} history of our Milky Way, about $2 \unit{Gyr}$ ago. This reflects the fact that binaries in our simulation typically take about a Gyr to merge.

When comparing the distribution of formation times for the three different DCO types we see that NSNSs are most strongly concentrated at recent times, followed by BHBHs and then BHNSs. To understand this it is helpful to consider that the inspiral time scales with various parameters
\begin{equation}
    t_{\rm inspiral} \propto \frac{a^4}{(m_1 + m_2)^{3}} \cdot \frac{q}{(1+q)^2} \cdot (1 - e^2)^{-7/2}.
\end{equation}
The inspiral time depends most strongly on the separation at DCO formation ($a$) and this is where the three types also differ most strongly (see Fig.~\ref{fig:dco_formation_properties}). The NSNS systems have the tightest orbits at DCO formation. The median of the distribution of separations at DCO formation, $\langle a_{\rm DCO} \rangle_{\rm med}$, relate as 8:3:1 for BHBH:BHNS:NSNS in our simulations. This results in increase of the inspiral time by a factor of about 4000:80:1. The total masses affect the inspiral time to the third power and this where the heavier BHBH systems are favoured. The median total masses differ by ratios of 6:4:1 for BHBH:BHNS:NSNS in our simulations, impacting the inspiral times such that they are a factor of 200:60:1 shorter, partially counteracting the effect of the separations. The term depending on the mass ratio $q$ only varies by about 30\% for the mass ratio ranges considered here and so is not of interest. The eccentricity term is not of importance for mildly eccentric systems, $f(e_{\rm DCO} \le 0.3)  \le 1.4$ but of large importance for the very eccentric  $f(e_{\rm DCO} \ge 0.9) \ge 300$. The fraction of highly eccentric systems with $e_{\rm DCO}>0.9$ is 33\%, 16\%, and 8\% of for BHBH, BHNS and NSNS respectively, see also Fig.~\ref{fig:dco_formation_properties}.

@@

\paragraph{Time until merger} Fig.~\ref{fig:fiducial_pdf_distributions}f shows the remaining time until merger for each of the DCO types at the start of LISA mission. The distributions are strikingly similar and peak with merger times of around one Myr.

@@

\subsection{Distribution in the Milky Way}\label{sec:mw_detectable_distribution}
For our Milky way model we considered three different components, a low-\achem~(``thin'') disc, an older high-\achem~(``thick'') disc and a bulge/bar. In Table~\ref{tab:component_fractions} we summarise the number of detections originating from each of these components. Despite the fact that only $42.5\%$ of systems are formed in the low-\achem~disc, we find that $86\%$ of the detections originate from this component. This is because most detectable systems were formed relatively recently (see Fig.~\ref{fig:fiducial_pdf_distributions}e) and so the high-\achem~disc and bulge are effectively too old to contribute many detectable systems. Nevertheless, we do find a significant fraction of detectable systems originate in the high-\achem~disc and bulge, indicating that it is still important to include these components, as was ignored in some earlier works (see Sec.~\ref{sec:compare_studies}).

\begin{table}[tb]
    \centering
    \begin{tabular}{l|ccccc}
        \hline
        \multirow{2}{*}{Component} & Formation & \multicolumn{4}{c}{Detectable} \\ \cline{2-6}
        & {\scriptsize All} & {\scriptsize All} & {\scriptsize BHBH} & {\scriptsize BHNS} & {\scriptsize NSNS} \\
        \hline
        Low-$\alpha$ disc & 42.5\% & 86\% & 89\% & 82\% & 85\%\\
        High-$\alpha$ disc & 42.5\% & 6\% & 5\% & 8\% & 12\%\\
        Bulge/bar & 15.0\% & 8\% & 6\% & 10\% & 3\%\\
    \end{tabular}
    \caption{Percentage of systems in each Galactic component.}
    \label{tab:component_fractions}
\end{table}
\begin{figure}[tb]
    \centering
    \includegraphics[width=\columnwidth]{fig4_1_galaxy_density_distribution.png}
    \includegraphics[width=\columnwidth]{fig4_2_detectable_distance_distribution_4yr.pdf}
    \caption{
    \textbf{Top:} A face-on view of the Galactic density distribution for detectable DCOs. We show the density distribution for the top 95\% of the sources, the rest are indicated by scatter points whose sizes correspond to their sampling weights. The coloured lines show the 75\% contour for each of the individual DCO types. The large cross passes through (0, 0) and helps to highlight the bias towards the position of the sun, which is indicated by the $\odot$.
    \textbf{Bottom:} As Figure~\ref{fig:fiducial_pdf_distributions}, but for the luminosity distance. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig4_1_galaxy_density_distribution.png}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/galaxy_creation_station.ipynb}{\faBook} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig4_2_detectable_distance_distribution_4yr.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/fiducial.ipynb}{\faBook}.}
    \label{fig:detectable_distance_dist}
\end{figure}

@@

The reason that the NSNSs are favoured at higher distances is that the NSNS population has the highest fraction of ``mildly'' eccentric systems ($0.01 < e < 0.03$). In contrast, the BHNS population has a much higher fraction of effectively circular systems ($e < 0.01$), which emit weaker gravitational wave compared to equivalent eccentric systems. Therefore, despite their typically higher masses, the distance at which a BHNS source is detectable is generally lower than for the mildly eccentric NSNS.

@@

\subsection{Formation channels}\label{sec:progenitors_and_formation}
In our fiducial model, approximately two thirds of detectable BHBHs are formed through the `only stable mass transfer' channel \citep{vandenHeuvel+2017}, whilst the remaining third are primarily formed through the `classic' CE channel \citep{Bhattacharya+1991}. Detectable BHNSs follow an inverse pattern, such that around two thirds are formed through the classic channel and the rest are mainly formed through only stable mass transfer.

@@

\subsection{How accurately will we be able to infer the parameters of detected systems?}\label{sec:measurement_uncertainties}
% So far we have discussed the properties of the detectable sources that can, in principle, be inferred from observations. However, many sources will only be barely detectable. Here we discuss for how many systems we expect to get high enough SNR to obtain accurate and useful measurements of these parameters. Below 
Here we discuss the expected SNR distribution and the typical uncertainties expected for the most relevant parameters, namely, the chirp mass and sky localisation.

@@

In Fig.~\ref{fig:snr_dist} we show the cumulative number of detections with a given SNR. Although many LISA sources have SNRs around the our assumed detection threshold of 7, many systems are detection with very high SNRs. We find that for a 4(10)-year LISA mission, of the 124(202) detections, 85(138), 16(27), and 9(14) systems have SNRs greater than 10, 50 and 100, respectively. These high SNR systems are typically, but not only, the more massive BHBH systems.

@@

We find that for a 4(10)-year LISA mission, approximately 41(105) detections have chirp masses ($\Delta \mathcal{M}_c / \mathcal{M}_c < 1$, indicated by the dark shaded region in Fig.~\ref{fig:m_c_unc}) whilst 13(31) have chirp mass uncertainties below 10\%. This uncertainty is generally dominated by the uncertainty on the time derivative of the frequency, since most of the binaries are too early in their inspiral for LISA to measure a strong chirp.
%
Note that increasing the mission length significantly increases the number of detections for which we can measure the mass. The total number of detections only scales as $\sqrt{T_{\rm obs}}$, yet we find that the total number of detections \textit{with well measured masses} scales approximately as $T_{\rm obs}$.

@@

We quantify the sky localisation of a source by estimating the angular resolution for the detectable sources. Since all potential sources are effectively stationary on the timescale of the LISA mission, we can follow \citet{Mandel+2018}, and use the timing accuracy of LISA and the effective detector baseline to calculate the angular resolution, $\sigma_{\theta}$, as
\begin{equation}
    \sigma_{\theta} = 16.6^\circ \left(\frac{7}{\rho}\right) \left(\frac{5 \times 10^{-4} \unit{Hz}}{f_{\rm dom}}\right) \left( \frac{2 \unit{AU}}{L} \right),
\end{equation}
where $L$ is the effective detector baseline, which for LISA is 2 AU since it will orbit the Sun.

>>>>

In this section we present our predictions for the population of detectable LISA sources for our fiducial model. In total we expect, on average, $124$ detections in a 4-year LISA mission, of which $\BHBHFourYear{}$, $\BHNSFourYear{}$ and $\NSNSFourYear{}$ are BHBHs, BHNSs and NSNSs respectively, based on our fiducial simulations. In the remainder of this section, we discuss where the sources are expected relative to LISA's sensitivity curve (Sec.~\ref{sec:dcos_on_sc}), their properties (Sec.~\ref{sec:fiducial_distributions}), their locations in the Milky Way (Sec.~\ref{sec:mw_detectable_distribution}), their formation channels (Sec.~\ref{sec:progenitors_and_formation}) and finally we discuss the expected SNR and how accurately we expect that the parameters can be measured (Sec.~\ref{sec:measurement_uncertainties}). Note that all results shown in this section are based on our fiducial simulations. A discussion of the impact of variations in the physics assumptions is provided in Sec.~\ref{sec:variations}.

@@

For both the BHBHs and BHNSs, we see that the black hole mass distribution favours low masses. About $90\%$ of BHs have masses below $11 \unit{M_{\odot}}$, in our fiducial simulations shown here. This is in stark contrast with observations from ground-based GW detectors, where heavy BHs with masses of about $30\unit{M_{\odot}}$ and higher have been common. There are two main reasons for this discrepancy. First, the population of BHs in the Milky Way (and, in particular, those detectable by LISA due to their recent formation times, see Figure~\ref{fig:fiducial_pdf_distributions}e) primarily come from progenitors that formed from high metallicity gas according to our simulations. Stellar winds are stronger at high metallicity leading to increased mass loss. This affects the mass of the most massive black holes that can be formed \citep{Belczynski+2010}. Secondly, ground-based detectors are strongly biased towards high mass systems, since they can be detected out to larger distances and thus a greater volume is probed. In contrast, LISA has no such bias and is, in principle, sensitive to inspiraling BHBHs throughout the entire Milky Way regardless of their mass, as long as we catch them when they are emitting in the LISA band. For this reason, LISA is more likely to detect the more common lower mass BHBHs.

We also note that our BH mass distribution extends down below $5\unit{M_{\odot}}$ to $2.5 \unit{M_{\odot}}$ which is our fiducial maximum neutron star mass. The BHs in this simulation fill the so-called ``lower mass gap'' marked as a grey band \citep{Ozel+2010,Farr+2011}, see also \citet{Shao+2021} who recently pointed this out. This prediction is sensitive to adopted model for fallback during the SN explosion as we discuss this further in Section~\ref{sec:lower_mass_gap}

\paragraph{Mass Ratio} The mass ratio distributions for each DCO type are very distinct from one another, as can be seen in Fig.~\ref{fig:fiducial_pdf_distributions}c. The majority of NSNSs have a mass ratio close to unity, with $\NSNSqAbovePointEight{}$ of systems having $q > 0.8$, where $q \equiv m_2 / m_1$. The reason for the concentration around equal masses is that most NSs are formed either through electron-capture supernovae (ECSN) or from low mass stars in our simulations. We assume a remnant mass for any NS formed through ECSN of $1.26 \unit{M_{\odot}}$ (see Sec.~\ref{app:fiducial_physics}). The remnant mass prescription that we use assumes a fixed fallback mass for any star with a CO core mass less than $2.5 \unit{M_\odot}$, such that many NSs end up with an identical mass of $1.278 \unit{M_\odot}$ \citep[see][Eq.~19]{Fryer+2012}. This means that many NSs are formed with equal masses and hence we see a mass ratio distribution peaked around unity.

@@

At the same time, stellar wind mass loss disfavours the formation of black holes with comparable masses. Such systems would have originated from progenitors that also started with comparable masses. The rather massive secondaries in these systems (especially after they accreted from the primary) experience very strong stellar wind mass loss due to LBV-like eruptions. This limits the amount by which they can expand \citep[e.g.][]{vanSon+2021}. At the same time, wind mass loss (and also stable and more conservative mas transfer in these systems) lead to widening of the orbit. Both effects, the reduced expansion and increased widening of the orbit, tend to prevent the secondaries from being able to fill their Roche lobe. This thus limits the number of systems that experience the reverse mass transfer or common-envelope phase needed shrink the orbits and make them tight enough to be detected as gravitational-wave sources.

We find that detectable BHNSs have even more unequal mass ratios. Moreover, the mass ratio distribution is bimodal, where the two peaks arise from two distinct formation scenarios. Around two thirds of detectable BHNSs experience at least one common-envelope event, whilst the last third are formed through only stable mass transfer. The first peak at $q = 0.18$ is from systems that experience at least one common-envelope phase and occurs at the expected mass ratio, which approximately follows the mean BH mass ($\sim 6.5 \unit{M_\odot}$) and NS mass ($\sim 1.2 \unit{M_\odot}$). Yet we also see a second peak at higher mass ratios around $q = 0.34$, which arises from the fraction of the population that underwent only stable mass transfer phases. The stability of mass transfer depends on the mass ratio, preferentially forming systems with more equal masses, i.e.\ at higher $q$.

\paragraph{Eccentricity} In Fig.~\ref{fig:fiducial_pdf_distributions}d we show the eccentricity distributions. We find that most systems (73\%) will have eccentricities larger than $0.01$ during the LISA mission, which should in principle be detectable according to \citet{Nishizawa+2016}. This means that we will potentially be able to use eccentricity to distinguish these sources from WDWDs, which are expected to have little to no eccentricity (see Sec.~\ref{sec:WDWD_distinguish}). We note that several previous studies assumed all systems to be circular when calculating the detection rates \citep[e.g.][]{Liu+2014, Lamberts+2018, Sesana+2020}. We discuss the impact of this assumption in Section~\ref{sec:compare_studies}.

For systems with eccentricities higher than $e\gtrsim0.3$, most gravitational wave energy is emitted in higher harmonics. Such systems are more rare, but we find them to be significant among the BHBH population, where they account for \BHBHHighlyEccentric{} of systems. Detectable BHBHs in our simulation (and, in particular, those that are eccentric) are primarily systems formed through the stable mass transfer channel (see Section~\ref{fig:formation_channels}). These systems are still relatively wide (compared to those formed through the CE channel) immediately prior to formation of the second BH, which makes them more easily affected by kicks. If the kick is oriented roughly in opposite direction to the orbital motion and has a velocity that is of similar magnitude as the orbital velocity, it will lead to the formation of a highly eccentric system.

It is rare to get such a ``lucky kick'', but there are a few effects that favour this for BHBHs. The kicks of BHs are reduced by fallback and they are thus less likely to disrupt the system. Moreover, because of their higher masses, BHBHs can be observed already at lower orbital frequencies. This means that they have not had as much time to circularise and so still have significant eccentricity by the time of the LISA mission. Finally, LISA favours the detection of eccentric systems, if all other properties are held fixed. This is because the gravitational-wave emission is stronger (Eq.~\ref{eq:peters_f}) and the energy is emitted at higher frequencies \citep[Eq.~20]{Peters+1963} where LISA is more sensitive.

@@

\paragraph{Time since formation} 
In Fig.~\ref{fig:fiducial_pdf_distributions}e we show how long ago the LISA detectable DCOs formed. Star formation was highest at early times $6-12\unit{Gyr}$ ago, after which it declined. In contrast, the LISA detectable DCOs primarily formed in the relatively \textit{recent} history of our Milky Way, about $2 \unit{Gyr}$ ago. This reflects the fact that binaries in our simulation typically take about a Gyr to merge.

When comparing the distribution of formation times for the three different DCO types we see that NSNSs are most strongly concentrated at recent times, followed by BHBHs and then BHNSs. To understand this it is helpful to consider how the inspiral time scales with various parameters \citep{Peters+1964}
\begin{equation}
    t_{\rm inspiral} \propto \frac{a^4}{(m_1 + m_2)^{3}} \cdot \frac{q}{(1+q)^2} \cdot (1 - e^2)^{-7/2}.
\end{equation}
The inspiral time depends most strongly on the separation at DCO formation ($a$) and this is where the three types also differ most strongly (see Fig.~\ref{fig:dco_formation_properties}). The detectable NSNS systems have the tightest orbits at DCO formation. The median of the distribution of separations at DCO formation, $\langle a_{\rm DCO} \rangle_{\rm med}$, relate as 8:3:1 for detectable BHBH:BHNS:NSNS in our simulations. This results in increase of the inspiral time by a factor of about 4000:80:1. The total masses affect the inspiral time to the third power and this where the heavier BHBH systems are favoured. The median total masses differ by ratios of 6:4:1 for detectable BHBH:BHNS:NSNS in our simulations, impacting the inspiral times such that they are a factor of 200:60:1 shorter, partially counteracting the effect of the separations. The term depending on the mass ratio $q$ only varies by about 30\% for the mass ratio ranges considered here and so is not of interest. The eccentricity term is not of importance for mildly eccentric systems, $f(e_{\rm DCO} \le 0.3)  \le 1.4$ but of large importance for the very eccentric  $f(e_{\rm DCO} \ge 0.9) \ge 300$. The fraction of highly eccentric systems with $e_{\rm DCO}>0.9$ is 33\%, 16\%, and 8\% of for BHBH, BHNS and NSNS respectively, see also Fig.~\ref{fig:dco_formation_properties}.

@@

\paragraph{Time until merger} Fig.~\ref{fig:fiducial_pdf_distributions}f shows the remaining time until merger for each of the DCO types at the start of LISA mission. The distributions are strikingly similar and peak with merger times of around a Myr.

@@

\subsection{Distribution in the Milky Way}\label{sec:mw_detectable_distribution}
For our Milky way model we consider three different components, a low-\achem~(``thin'') disc, an older high-\achem~(``thick'') disc and a bulge/bar. In Table~\ref{tab:component_fractions} we summarise the number of detections originating from each of these components. Despite the fact that only $42.5\%$ of systems are formed in the low-\achem~disc, we find that $86\%$ of the detections originate from this component. This is because most detectable systems were formed relatively recently (see Fig.~\ref{fig:fiducial_pdf_distributions}e) and so the high-\achem~disc and bulge are effectively too old to contribute many detectable systems. Nevertheless, we do find a significant fraction of detectable systems originate in the high-\achem~disc and bulge, indicating that it is still important to include these components, as was ignored in some earlier works (see Sec.~\ref{sec:compare_studies}).

\begin{table}[tb]
    \centering
    \begin{tabular}{l|c|cccc}
        \hline
        \multirow{2}{*}{Component} & Formation & \multicolumn{4}{c}{Detectable} \\ \cline{2-6}
        & {\scriptsize All} & {\scriptsize All} & {\scriptsize BHBH} & {\scriptsize BHNS} & {\scriptsize NSNS} \\
        \hline
        Low-$\alpha$ disc & 42.5\% & 86\% & 89\% & 82\% & 85\%\\
        High-$\alpha$ disc & 42.5\% & 6\% & 5\% & 8\% & 12\%\\
        Bulge/bar & 15.0\% & 8\% & 6\% & 10\% & 3\%\\
    \end{tabular}
    \caption{Percentage of systems in each Galactic component.}
    \label{tab:component_fractions}
\end{table}
\begin{figure}[tb]
    \centering
    \includegraphics[width=\columnwidth]{fig4_1_galaxy_density_distribution.png}
    \includegraphics[width=\columnwidth]{fig4_2_detectable_distance_distribution_4yr.pdf}
    \caption{
    \textbf{Top:} A face-on view of the Galactic density distribution for detectable DCOs. We show the density distribution for the top 95\% of the sources, the rest are indicated by scatter points whose sizes correspond to their sampling weights. The coloured lines show the 75\% contour for each of the individual DCO types. The large cross passes through (0, 0) and helps to highlight the bias towards the position of the sun, which is indicated by the $\odot$.
    \textbf{Bottom:} As Figure~\ref{fig:fiducial_pdf_distributions}, but for the luminosity distance. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig4_1_galaxy_density_distribution.png}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/galaxy_creation_station.ipynb}{\faBook} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig4_2_detectable_distance_distribution_4yr.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/fiducial.ipynb}{\faBook}.}
    \label{fig:detectable_distance_dist}
\end{figure}

@@

The reason that the NSNSs are favoured at higher distances is that the NSNS population has the highest fraction of ``mildly'' eccentric systems ($0.01 < e < 0.03$). In contrast, the BHNS population has a much higher fraction of effectively circular systems ($e < 0.01$), which emit weaker gravitational waves compared to equivalent eccentric systems. Therefore, despite their typically higher masses, the distance at which a BHNS source is detectable is generally lower than for the mildly eccentric NSNS.

@@

\subsection{Formation channels}\label{sec:progenitors_and_formation}
In our fiducial model, approximately two thirds of detectable BHBHs are formed through the `only stable mass transfer' channel, whilst the remaining third are primarily formed through the `classic' CE channel. Detectable BHNSs follow an inverse pattern, such that around two thirds are formed through the classic channel and the rest are mainly formed through only stable mass transfer.

@@

\subsection{How accurately will we be able to infer the parameters of detected systems?}\label{sec:measurement_uncertainties}
So far we have discussed the properties of the detectable sources that can, in principle, be inferred from observations. However, many sources will only be barely detectable. Here we discuss for how many systems we expect to get high enough SNR to obtain accurate and useful measurements of these parameters. Below we discuss the expected SNR distribution and the typical uncertainties expected for the most relevant parameters, namely, the chirp mass and sky localisation.

@@

In Fig.~\ref{fig:snr_dist} we show the cumulative number of detections with a given SNR. Although many a large fraction of sources have SNRs around our assumed detection threshold of 7, many systems are detected with very high SNRs. We find that on average for a 4(10)-year LISA mission, of the 124(202) detections, 85(138), 16(27) and 9(14) systems have SNRs greater than 10, 50 and 100, respectively. These high SNR systems are typically, but not only, the more massive BHBH systems.

@@

We find that for a 4(10)-year LISA mission, approximately 41(105) detections have measurable chirp masses ($\Delta \mathcal{M}_c / \mathcal{M}_c < 1$, indicated by the dark shaded region in Fig.~\ref{fig:m_c_unc}) whilst 13(31) have chirp mass uncertainties below 10\% ($\Delta \mathcal{M}_c / \mathcal{M}_c < 0.1$, indicated by the light shaded region in Fig.~\ref{fig:m_c_unc}). This uncertainty is generally dominated by the uncertainty on the time derivative of the frequency, since most of the binaries are too early in their inspiral for LISA to measure a strong chirp.
%
Note from Fig.~\ref{fig:m_c_unc} that increasing the mission length significantly increases the number of detections for which the chirp mass uncertainty is below 100\%. The total number of detections only scales as $\sqrt{T_{\rm obs}}$, yet we find that the the total number of detections with a chirp mass uncertainty is below 100\% and 10\% both scale approximately as $T_{\rm obs}$.

@@

We quantify the sky localisation of a source by estimating the angular resolution for the detectable sources. Since all potential sources are effectively stationary on the timescale of the LISA mission, we can follow \citet{Mandel+2018} and use the timing accuracy of LISA and the effective detector baseline to calculate the angular resolution, $\sigma_{\theta}$, as
\begin{equation}
    \sigma_{\theta} = 16.6^\circ \left(\frac{7}{\rho}\right) \left(\frac{5 \times 10^{-4} \unit{Hz}}{f_{\rm dom}}\right) \left( \frac{2 \unit{AU}}{L} \right),
\end{equation}
where $L$ is the effective detector baseline, which for LISA is 2 AU in the frame of the solar system, since it will orbit the Sun.
